R Notebook

### 2019-10-22 src/testing_file.Rmd
fandangled plotting

boxplot of test auc scores test

@@

boxplot of train auc scores test

@@

boxplot of cbi

@@

boxplots of tss scores
```{r}
dec <- read.csv("../output/maxent/models/decavgAIC.csv", header = TRUE)
mar <- read.csv("../output/maxent/models/maravgAIC.csv", header = TRUE)
apr <- read.csv("../output/maxent/models/apravgAIC.csv", header = TRUE)
may <- read.csv("../output/maxent/models/mayavgAIC.csv", header = TRUE)
jun <- read.csv("../output/maxent/models/junavgAIC.csv", header = TRUE)
jul <- read.csv("../output/maxent/models/julavgAIC.csv", header = TRUE)
aug <- read.csv("../output/maxent/models/augavgAIC.csv", header = TRUE)
sep <- read.csv("../output/maxent/models/sepavgAIC.csv", header = TRUE)
oct <- read.csv("../output/maxent/models/octavgAIC.csv", header = TRUE)
nov <- read.csv("../output/maxent/models/novavgAIC.csv", header = TRUE)

@@

for (m in 1:length(mthord)){
  mthna <- mthord[m]
  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*asc", full.names = TRUE)
  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    ascname <- gsub("avg_cloglog_", "", ascname)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna,  "_binary_0.5_", ascname, ".asc"), overwrite = TRUE)
     plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_binary_0.5_", ascname, ".png")), overwrite = TRUE)
       dev.off() # stops automatic saving of the plot to a png
  }
}

@@

for (m in 1:length(mthord)){
  mthna <- mthord[m]
  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_binary_0.5_.*.asc$"), full.names = TRUE)
  binstack <- stack(binaryprob_list)
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.asc"), overwrite=TRUE)
  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.png")))
  dev.off()
}

@@

aea <- raster("../output/env/aea.tif") #the correct CRS
maplst <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "_bin_0.5_2d.*.asc", full.names = TRUE)
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
apr <- aea@crs
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")

@@

# add seabed?
need to add seabed

@@

>>>>

#boxplot of test auc scores test

@@

#boxplot of train auc scores test

@@

#boxplot of cbi

@@

#boxplots of tss scores
```{r}
dec <- read.csv("../output/maxent/models/decavgAIC.csv", header = TRUE)
mar <- read.csv("../output/maxent/models/maravgAIC.csv", header = TRUE)
apr <- read.csv("../output/maxent/models/apravgAIC.csv", header = TRUE)
may <- read.csv("../output/maxent/models/mayavgAIC.csv", header = TRUE)
jun <- read.csv("../output/maxent/models/junavgAIC.csv", header = TRUE)
jul <- read.csv("../output/maxent/models/julavgAIC.csv", header = TRUE)
aug <- read.csv("../output/maxent/models/augavgAIC.csv", header = TRUE)
sep <- read.csv("../output/maxent/models/sepavgAIC.csv", header = TRUE)
oct <- read.csv("../output/maxent/models/octavgAIC.csv", header = TRUE)
nov <- read.csv("../output/maxent/models/novavgAIC.csv", header = TRUE)

@@

for (m in 1:length(mthord)){
  mthna <- mthord[m]
  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*asc", full.names = TRUE)
  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    ascname <- gsub("avg_cloglog_", "", ascname)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/predictions/mar/", mthna,  "_binary_0.5_", ascname, ".asc"), overwrite = TRUE)
     plot(probbin, png(paste0("../output/maxent/predictions/mar/", mthna, "_binary_0.5_", ascname, ".png")), overwrite = TRUE)
       dev.off() # stops automatic saving of the plot to a png
  }
}

@@

for (m in 1:length(mthord)){
  mthna <- mthord[m]
  binaryprob_list <- list.files(paste0("../output/maxent/predictions/mar/"), pattern = paste0(mthna, "*_binary_0.5_.*.asc$"), full.names = TRUE)
  binstack <- stack(binaryprob_list)
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/predictions/mar/", mthna, "_bin_0.5_2d.asc"), overwrite=TRUE)
  plot(stacksum, png(paste0("../output/maxent/predictions/mar/", mthna, "_bin_0.5_2d.png")))
  dev.off()
}

@@

aea <- raster("../output/env/aea.tif") #the correct CRS
maplst <- list.files("../output/maxent/predictions/mar/", pattern = "_bin_0.5_2d.*.asc", full.names = TRUE)
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
apr <- aea@crs
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")

@@

# add contour?
need to add seabed

@@

#coloured histograms

```{r}
mar <- raster("../output/maxent/predictions/mar/avg_cloglog_97.0413.asc")
#apr <- raster("../output/maxent/predictions/apr/avg_cloglog_97.0413.asc")

mardf <- as.data.frame(mar) #convert raster to dataframe
mardf <- na.omit(mardf) #remove all the land 
for (i in 1:nrow(mardf)){
  if (mardf$avg_cloglog_97.0413[i] <=0.1){
    mardf$bin[i] <- 1
  } else if (mardf$avg_cloglog_97.0413[i] >0.1 & mardf$avg_cloglog_97.0413[i] <= 0.2){
      mardf$bin[i] <- 2
  } else if (mardf$avg_cloglog_97.0413[i] >0.2 & mardf$avg_cloglog_97.0413[i] <= 0.3){
      mardf$bin[i] <- 3
  } else if (mardf$avg_cloglog_97.0413[i] >0.3 & mardf$avg_cloglog_97.0413[i] <= 0.4){
      mardf$bin[i] <- 4
  } else if (mardf$avg_cloglog_97.0413[i] >0.4 & mardf$avg_cloglog_97.0413[i] <= 0.5){
      mardf$bin[i] <- 5
  } else if (mardf$avg_cloglog_97.0413[i] >0.5 & mardf$avg_cloglog_97.0413[i] <= 0.6){
      mardf$bin[i] <- 6
  } else if (mardf$avg_cloglog_97.0413[i] >0.6 & mardf$avg_cloglog_97.0413[i] <= 0.7){
      mardf$bin[i] <- 7
  } else if (mardf$avg_cloglog_97.0413[i] >0.7 & mardf$avg_cloglog_97.0413[i] <= 0.8){
      mardf$bin[i] <- 8
  } else if (mardf$avg_cloglog_97.0413[i] >0.8 & mardf$avg_cloglog_97.0413[i] <= 0.9){
      mardf$bin[i] <- 9
  } else if (mardf$avg_cloglog_97.0413[i] >0.9){
      mardf$bin[i] <- 10
  }
}

mardf$bin <- as.factor(mardf$bin)

brk <- seq(0.1, 1, 0.1) #bins you want
brkl <- as.character(brk)

p <- ggplot(mardf, aes(x = avg_cloglog_97.0413, fill = bin))+ geom_histogram(breaks = brk) + theme_classic() + scale_fill_viridis_d(option = "inferno", direction = -1) + theme(legend.position = "none", plot.title = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank()) + scale_x_continuous(breaks=seq(0,1,0.1))
png(paste0("../output/maxent/paper_plots/histogram_", mthna, "_", dep, ".png"), height = 1024, width = 1024)
p
dev.off()

#selected maps pt2

get unique depths
```{r}
mar <- read.csv("../output//maxent/paper_plots/prob_by_depth/mar_u0.4s_dis_prob_cells.csv", header = TRUE)
mar <- mar[!(mar$Depth =="1151.99" | mar$Depth =="1265.86" | mar$Depth =="1387.38"), ]
mar$Probability <- gsub(">0.9", "0.9 - 1", mar$Probability)
mar$Depth <- as.factor(mar$Depth)
flevelsm <- levels(mar$Depth)
```

aea <- raster("../output/env/aea.tif") #the correct CRS
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
acrs <- aea@crs
mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
ss <- raster::intersect(s, mcp)  #trim land file to mcp

#selmth <- c("mar", "jun", "sep", "dec")
#seldep <- c("11.7737", "61.1128", "199.79", "628.026")

mar11 <- raster("../output/maxent/predictions/mar/avg_cloglog_11.7737.asc")
mar61 <- raster("../output/maxent/predictions/mar/avg_cloglog_61.1128.asc")
mar199 <- raster("../output/maxent/predictions/mar/avg_cloglog_199.79.asc")
mar628 <- raster("../output/maxent/predictions/mar/avg_cloglog_628.026.asc")
mar11@crs <- acrs
mar61@crs <- acrs
mar199@crs <- acrs
mar628@crs <- acrs
mar11 <- raster::mask(mar11, mcp)
mar61 <- raster::mask(mar61, mcp)
mar199 <- raster::mask(mar199, mcp)
mar628 <- raster::mask(mar628, mcp)

jun11 <- raster("../output/maxent/predictions/jun/avg_cloglog_11.7737.asc")
jun61 <- raster("../output/maxent/predictions/jun/avg_cloglog_61.1128.asc")
jun199 <- raster("../output/maxent/predictions/jun/avg_cloglog_199.79.asc")
jun628 <- raster("../output/maxent/predictions/jun/avg_cloglog_628.026.asc")
jun11@crs <- acrs
jun61@crs <- acrs
jun199@crs <- acrs
jun628@crs <- acrs
jun11 <- raster::mask(jun11, mcp)
jun61 <- raster::mask(jun61, mcp)
jun199 <- raster::mask(jun199, mcp)
jun628 <- raster::mask(jun628, mcp)

sep11 <- raster("../output/maxent/predictions/sep/avg_cloglog_11.7737.asc")
sep61 <- raster("../output/maxent/predictions/sep/avg_cloglog_61.1128.asc")
sep199 <- raster("../output/maxent/predictions/sep/avg_cloglog_199.79.asc")
sep628 <- raster("../output/maxent/predictions/sep/avg_cloglog_628.026.asc")
sep11@crs <- acrs
sep61@crs <- acrs
sep199@crs <- acrs
sep628@crs <- acrs
sep11 <- raster::mask(sep11, mcp)
sep61 <- raster::mask(sep61, mcp)
sep199 <- raster::mask(sep199, mcp)
sep628 <- raster::mask(sep628, mcp)

dec11 <- raster("../output/maxent/predictions/dec/avg_cloglog_11.7737.asc")
dec61 <- raster("../output/maxent/predictions/dec/avg_cloglog_61.1128.asc")
dec199 <- raster("../output/maxent/predictions/dec/avg_cloglog_199.79.asc")
dec628 <- raster("../output/maxent/predictions/dec/avg_cloglog_628.026.asc")
dec11@crs <- acrs
dec61@crs <- acrs
dec199@crs <- acrs
dec628@crs <- acrs
dec11 <- raster::mask(dec11, mcp)
dec61 <- raster::mask(dec61, mcp)
dec199 <- raster::mask(dec199, mcp)
dec628 <- raster::mask(dec628, mcp)

#2d plots
mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
mar2d <- raster::mask(mar2d, mcp)
mar2d@crs <- acrs
jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
jun2d <- raster::mask(jun2d, mcp)
jun2d@crs <- acrs
sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
sep2d <- raster::mask(sep2d, mcp)
sep2d@crs <- acrs
dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")
dec2d <- raster::mask(dec2d, mcp)
dec2d@crs <- acrs

#mar2d, jun2d, sep2d, dec2d,

pstack <- stack(mar11, jun11, sep11, dec11, mar61, jun61, sep61, dec61, mar199, jun199, sep199, dec199, mar628, jun628, sep628, dec628) #stack the layers - printed row-wise

rws <- c("11.7737 m", "61.1128 m", "199.70 m", "628.026 m", "", "", "", "", "", "", "", "", "", "", "", "")
cls <- c("March", "June", "September", "December", "", "", "", "", "", "", "", "", "", "", "", "")
myTheme <- rasterTheme(region = inferno(rev(10)))
my.at <- seq(0, 1, by = 0.1)

p <- levelplot(pstack, ylab = NULL, xlab = NULL, colorkey=list(space="bottom", at = my.at), par.settings = myTheme, layout = c(4, 4), margin = FALSE, scales=list(draw=FALSE), col.regions=rev(inferno(10)), at= my.at, names.attr = cls) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
png("../output/maxent/paper_plots/bydepth.png", height = 1024, width = 1024)
p
dev.off()

# variable maps ALSO FOR P2
- temp
- chl 
- o2
- sal
- 4 different months
- relationship of variables to each other
- pick a depth (say the 11.7737 m)

months are mar, jun, sep, dec
pick 3 diff years say 1998, 2006, 2013

```{r}
aea <- raster("../output/env/aea.tif") #the correct CRS
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
acrs <- aea@crs
mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
ss <- raster::intersect(s, mcp)  #trim land file to mcp

yrlst <- c("1998", "2006", "2013")
mthlst <- c("03", "06", "09", "12")
varlst <- c("temp", "salinity", "o2")

for (y in 1:length(yrlst)){
  yr <- yrlst[y]
  fdr <- paste0("../data/env/ascii/ascii", yr, "/")
  for (m in 1:length(mthlst)){
    mt <- mthlst[m]
    for (v in 1:length(varlst)){
      var <- varlst[v]
      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_11.7737.asc"))
      lyrnm <- names(ras)
      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      ras <- projectRaster(ras, aea)
      ras <- raster::mask(ras, mcp)

      p <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))

      png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
      print(p)
      dev.off()
    }
  }
}

#variable interactions... KEEP FOR P2

```{r}
yrlst <- c("1998", "2006", "2013")
mthlst <- c("03", "06", "09", "12")
varlst <- c("temp", "salinity", "o2")
scbr <- stack()

for (y in 1:length(yrlst)){
  yr <- yrlst[y]
  fdr <- paste0("../data/env/ascii/ascii", yr, "/")
  for (m in 1:length(mthlst)){
    mt <- mthlst[m]
    for (v in 1:length(varlst)){
      var <- varlst[v]
      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_11.7737.asc"))
      lyrnm <- names(ras)
      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      ras <- projectRaster(ras, aea)
      ras <- raster::mask(ras, mcp)
      scbr <- stack(scbr, ras)
    }
  }
}
a <- xyplot(X2013_12_temp.nc_11.7737~X2013_12_salinity.nc_11.7737, data = scbr, auto.key = list(space='right'))
b <- xyplot(X2013_06_temp.nc_11.7737~X2013_06_salinity.nc_11.7737, data = scbr, auto.key = list(space='right'))

a
b
```

```{r}
splom(scbr)
```

#2d plots - part 2

the ggplot experiment....

```{r}
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
acrs <- aea@crs
mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
ss <- raster::intersect(s, mcp)  #trim to mcp

mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
apr2d <- raster("../output/maxent/paper_plots/hotspot_mapping/apr_bin_0.5_2d_single.asc")
may2d <- raster("../output/maxent/paper_plots/hotspot_mapping/may_bin_0.5_2d_single.asc")
jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
jul2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jul_bin_0.5_2d_single.asc")
aug2d <- raster("../output/maxent/paper_plots/hotspot_mapping/aug_bin_0.5_2d_single.asc")
sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
oct2d <- raster("../output/maxent/paper_plots/hotspot_mapping/oct_bin_0.5_2d_single.asc")
nov2d <- raster("../output/maxent/paper_plots/hotspot_mapping/nov_bin_0.5_2d_single.asc")
dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")

#trim to mcp
mar2d <- raster::mask(mar2d, mcp)
apr2d <- raster::mask(apr2d, mcp)
may2d <- raster::mask(may2d, mcp)
jun2d <- raster::mask(jun2d, mcp)
jul2d <- raster::mask(jul2d, mcp)
aug2d <- raster::mask(aug2d, mcp)
sep2d <- raster::mask(sep2d, mcp)
oct2d <- raster::mask(oct2d, mcp)
nov2d <- raster::mask(nov2d, mcp)
dec2d <- raster::mask(dec2d, mcp)

#add crs
mar2d@crs <- acrs
apr2d@crs <- acrs
may2d@crs <- acrs
jun2d@crs <- acrs
jul2d@crs <- acrs
aug2d@crs <- acrs
sep2d@crs <- acrs
oct2d@crs <- acrs
nov2d@crs <- acrs
dec2d@crs <- acrs

ggplot() +  ggspatial::geom_spraster_rgb(mar2d) + coord_fixed()
```
